---
title       : Predict Car Mileage Web App
subtitle    : Feb 16, 2015
author      : Kanthimathi Gayatri Sukumar
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---

## Problem Statement

With raising gasoline costs and the call to conserve natural resources (in this case fossil fuel), it becomes very important to choose and use fuel efficient cars that provide a high mileage. 

In order to help users determine (predict) the mileage, for informed decision making, we have created a Web App that is capable of predicting the mileage given the cars number of cylinders, displacement, horsepower and weight with 83.27% variance explained.

The Web App is at https://g3murari.shinyapps.io/PredictMileage/

The Web App restricts the input values to the range that has been used for modelling.

--------------------------------------------------------------------------------

## Methods

The methods used for the prediction of mileage are - 

1. <b>Data Collection:</b> The dataset used for the prediction is the <b>mtcars</b> dataset available in R. As per R documentation, the data was extracted from the 1974 Motor Trend US magazine, and comprises fuel consumption and 10 aspects of automobile design and performance for 32 automobiles (1973-74 models).

2. <b>Exploratory Analysis:</b> Exploration of the data was performed using R, looking for covariance, zero variance and spread of the data.

3. <b> Statistical Modelling:</b> Parametric models like linear regression and non-parametric models like CART, Random Forest and Boosting were used. The model that has been able to explain variance the highest has been Random Forest with 83.27% variance explained.

--------------------------------------------------------------------------------

## Results

```{r loadLibs, echo=FALSE, results=FALSE, warning=FALSE, message=FALSE}
library (caret)
library (randomForest)
library (miscTools)
library (ggplot2)
data (mtcars)
set.seed (12345)
```

Below is the modelling for mileage on mtcars dataset using Random Forest algorithm with 3 fold cross validation. The results shows the 83.27% variance explained with an MSE of 5.48.

```{r randomForest}
fitControl = trainControl (method = "cv", number = 3)
rfFit = train (mpg ~ cyl + disp + hp + wt, data = mtcars, method = "rf", trControl = fitControl)
rfFit$finalModel
```


--------------------------------------------------------------------------------

## Conclusion

The plot of the Random Forest Fit R-squared shows high predictability and hence the Web App should be able to predict the mileage very close to the actual value.

```{r plotModel, echo=FALSE}
fitControl = trainControl (method = "cv", number = 3)
rfFit = train (mpg ~ cyl + disp + hp + wt, data = mtcars, method = "rf", trControl = fitControl)

pred = predict (rfFit, mtcars [,-1])
r2 = rSquared (mtcars$mpg, mtcars$mpg - pred)

p = ggplot (aes(x=actual, y=pred),
  data=data.frame(actual = mtcars$mpg, pred =  pred))

p + geom_point() +
    geom_abline(color="red") +
    ggtitle(paste("RandomForest Regression in R r^2=", r2, sep=""))
```


--------------------------------------------------------------------------------

## References

Dataset Reference: http://stat.ethz.ch/R-manual/R-devel/library/datasets/html/mtcars.html

Course Reference: https://class.coursera.org/devdataprod-011


